---
title: "Snowshoe Hares Report (ROZAL)"
author: "Sam Rozal"
date: "11/6/2021"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
#attach packages 
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(effsize)
library(kableExtra)
library(ggbeeswarm)
```

```{r, include=FALSE}
#read in the data 
hares <- read_csv(here("data", "bonanza_hares.csv"))
```

# 1. Introduction
This report provides an exploratory overview into differences in weights of juvenile snowshoe hares observed at the Bonanza Creek Long Term Ecological Research site (a portion of the US Long Term Ecological Research Network) between the years of 1998 and 2012. Specifically, I analyze the differences in juvenile snowshoe hare weights between sexes (male and female) and experimental sites (Bonanza Riparian, Bonanza Mature, and Bonanza Black Spruce). We also farther explore the relationship between juvenile hare weight and juvenile hare hind foot length by building a simple linear regression model is built to explore this relationship 

# 2. Data and Methods
Measurements pertaining to snowshoe hare sizes were collected by Dr. Knut Kielland, Dr. Staurt Chapin, Dr. Roger Russ, and their team between the years of 1998 and 2012 in the Bonanza Creek Experimental Forest in Alaska, USA. The data covers 5 sites (three of which are explored further in this report, as this report focuses on Juvenile snowshoe hares) where snowshoe hares were captured in traps, measured, and then released back into the the ecological reserve.Exploratory data visualization is first conducted in order to_____. Folowing this initial exploration, the difference between male and female weights is analysed using a Welch two-sample t-test, with a significance level of 0.05,the effect size of the model is found, and a linear regression model exploring weight and hind foot lengthis constructed.

```{r}
#Data organization 
hares_date <- hares %>% #create new df from  hares
  mutate(date_new = mdy(date)) %>% #use lubricate to create new colum with code-friendly data
  mutate(year = year(date_new)) #create year column using year() function 
```

```{r}
#sort data 
juvenile_hares_date <-hares_date %>% #create new data frame from hares_date
  filter(age == "j") %>% #filter for just juvinile hares
  mutate(sex_full = case_when(
    sex == "m" ~ "Male", 
    sex == "f" ~ "Female", 
    sex == "na" ~ "Not Applicable"
  )) %>% 
  mutate(site_full = case_when(
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonmat" ~ "Bonanza Mature", 
    grid == "bonbs" ~ "Bonanza Black Spruce"
  ))
```


```{r}
hares_summary <- juvenile_hares_date %>%  #create new df from hares_date
  group_by(year) %>% # group by year
  summarize(count=n()) #get count totals 
```

```{r}
#get summary stats
hares_summary_statistics <- hares_summary %>% 
  summarize(median= median(count),
            mean= mean(count))
```

# 3. Preliminary Results 
## A. Annual Juvenile Hare Trap Counts
```{r}
#Annual juvenile hare counts 
ggplot(data=hares_summary, aes(x=year, y=count)) +
  geom_col(fill="purple", color= "blue") + #create graph using geom_col
  theme_classic() +
  scale_x_continuous(breaks= seq(from=1998, to=2012, by=1)) + #makes it so that all years are shown 
  labs(x= "Year",
       y= "Juvenile Hares Trapped Count (n)", 
       title = "Number of Juvenile Hares Trapped by Year") + #add labels
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue"))

```

Figure 1.*Count of Juvenile Hares Trapped by year.* Bar values represent the total number of juvenile hares that were trapped and subsequently observed each year. The number of observed hares greatly varied from year to year with a maximum number of hares trapped (126) occurring in 1999, and the minimum number of hares trapped (2) occurring in 2010. Throughout the entire study period, there was a mean value of `r hares_summary_statistics$mean` hares observed and a median value of `r hares_summary_statistics$median` hares observed. These values, however, do not take into account effort as we do not______


## B.Juvenile hare weights 

```{r}
juvenile_hares_mean_weight <- juvenile_hares_date %>% 
  group_by(sex,grid) %>% 
  summarize(mean_weight= mean(weight, na.rm=TRUE))
```


```{r}
#Juvenile hare weights-final graph 
ggplot(data=juvenile_hares_date, aes(x=sex_full, y=weight))+
  geom_beeswarm(aes(color=sex_full))+
  labs(x="Site and Sex",
        y= "Juvenile Hare Weight (g)",
       title = "Juvenile Hare Weight by Sex and Site Trapped")+
  geom_boxplot(fill=NA, width= 0.4, outlier.color="red") +
  labs(color= "Sex")+
  theme(legend.background= element_rect(fill= "light blue"))+
  theme(plot.background = element_rect(fill= "azure")) +
  facet_wrap(~site_full) +
  stat_summary(fun=mean,
               geom="point",
               shape=18, 
               size=2)
       
```

**Figure.2 Juvenile Hare Weight by sex (male, female, or unidentified) and site trapped (Bonanza Black Spruce, Bonanza Mature, or Bonanza Riparian)** Pink, blue, and grey points represent individual trapped hare observations of female, male, and sex unidentified respectively. The solid black line within in the boxpolot represents the median number of hares trapped and the endpoints of the boxplot are representative of the 25th and 75th percentile. Average(mean) values of trapped hares are represented by black diamonds and outliers are represented by red points. Data: Kielland et al. (2017)

On averge, across all three sites, male juvenile snowshoe hares were larger than that of female juvenile snowshoe hares. In comparing the size of the shares across the three sites, hares that were trapped at the Bonanza Black Spruce site were largest. We see the greatest distribution of sizes among male hares that wer trapped at the Bonanza Riparian site.

## C.Comparison of Weights between Female and Male Snowshoe Hares

```{r}
juvenile_hares_weight <- juvenile_hares_date %>% #create a new data frame
  group_by(sex_full) %>%  #group by sex
  summarize(mean_hare_weight= round(mean(weight,na.rm=TRUE), 2),
            sd_hare_weight= round(sd(weight, na.rm=TRUE), 2), 
            hare_sample_size= n())#generate summary statistics

```
```{r} 
#Awesomely styled chart of juvenile weight comparison 
kable(juvenile_hares_weight, col.names= c("Sex", 
                                          "Mean Hare Weight (g)",
                                          "Standard Deviation(g)",
                                          "Sample Size (n)"), caption= "Table 1. Summary Statistics for Female, Male, and N/A Hares. Table includes mean, standard deviation, and sample size for Juvenile Hares of all sexes found at three diffent sites (Bonanza Riparian, Bonzana Mature, and Bonaza Black spruce) between the yeasr of 1998 and 2012") %>% 
  kable_styling()
```

```{r}
#Setting up for a potential T-test:
#create a vector with just female juvenile hares from the juvenile_hares_date df
juvenile_hares_female <- juvenile_hares_date %>% 
  filter(sex_full=="Female", weight != "NA") %>%
  pull(weight)

#create a vector with just male juvenile hares from the juvenile_hares_date df
juvenile_hares_male <- juvenile_hares_date %>% 
  filter(sex_full=="Male", weight !="NA") %>% 
  pull(weight)
  
```
```{r, include=FALSE}
#plot histogram and qq plots to check assumptions
hist(juvenile_hares_female) # pretty much normal, slight skew
qqnorm(juvenile_hares_female) #beautiful! Pretty linear, qq plot basically a straight line 
hist(juvenile_hares_male)# mostly normal distributed, slight skew bc there are a lot of hares in the 500 bin, which is a lot lower than it should be
qqnorm(juvenile_hares_male)# mostly linear, but there is a slight s curve

#Based on looking at this we can assume that the data is roughly normal, which means that we can in fact run a two sample t test. 
```
```{r}
# run the t.test:
juvenile_hares_t <- t.test(juvenile_hares_female,juvenile_hares_male) 
```
```{r}
#pull model summary of the data frame so that I can call the outputs for in line referencing 
juvenile_hares_t_tidy <- broom::tidy(juvenile_hares_t)
```

```{r}
#run a Cohen's effect test to get effect size 
juvenile_hares_cohens <- cohen.d(juvenile_hares_male,juvenile_hares_female)

#find the summary statistics in order to report it later on with in line referencing 
female_hares_mean <- mean(juvenile_hares_female)
female_hares_sd <- sd(juvenile_hares_female)
male_hares_mean <- mean(juvenile_hares_male)
male_hares_sd <- sd(juvenile_hares_male)
```

Our analysis indicates that male juvenile snowshoe hares have a larger average weight compared to that of the average weight of female juvenile snowshoe hares (`r round(male_hares_mean, 2)` $\pm$ `r round(male_hares_sd, 2)` and `r round(female_hares_mean, 2)` $\pm$ `r round(female_hares_sd, 2)` respectively, reported as mean $\pm$ 1 standard deviation). The absolute difference in means between females and males is `r round(male_hares_mean, 2) -  round(female_hares_mean, 2)` grams, which is a `r round((male_hares_mean - female_hares_mean)/((male_hares_mean + female_hares_mean)/2)*100, 2)` percent difference. Exploratory qq plots and histograms indicated a normal distribution of residuals and data respectively, so a two sided t-test was warranted to assess the significance of this difference. The difference between these means is significant (t(`r round(juvenile_hares_t_tidy$parameter, 2)`) = `r round(juvenile_hares_t_tidy$statistic, 2)`, p=`r juvenile_hares_t_tidy$p.value[1]`). This indicated that if drawing from populations with the same average value, there is a 0.71% chance that difference in mean values between female and male juvenile hares is a result of random error, which means that we can reject the null hypothesis that the true difference in means is not equal to zero.  This is corroborated by a small effects size (Cohen's *d* =`r round(juvenile_hares_cohens$estimate, 2)`) which indicates that ________. Data: Kielland et al. (2017)

## D. Relationship Between Weight and Hind Foot Length
We analyzed the relationship between juvenile hare weight (grams) and hind foot length (mm) across all juvenile hares observed during the study period. Although the relationship provides valuable insight , there is a necessity for further analysis that also includes sex of the juvenile hare and the site at which the hares were observed/trapped as variables, as these are likely to also have an impact on rabbit weight.
```{r}
#perform an linear regression analysis
juvenile_hares_lm <- lm(weight ~ hindft, data=juvenile_hares_date) #stored linear model as juvenile_hares_lm

#return tidy version of the model with the coefficient summary so that I can use it for in-line referencing later on.
juvenile_hares_lm_tidy <- broom::tidy(juvenile_hares_lm)

```

```{r}
#return the model summary to use for in line referencing 
juvenile_hares_lm_glance <- broom::glance(juvenile_hares_lm)
```
```{r}
#Code for Pearson's r correlation 
juvenile_hare_cor <- cor.test(juvenile_hares_date$hindft, juvenile_hares_date$weight)

#get tidy version of the correlation for use later on with in line referncing 
juvenile_hares_cor_tidy <- broom::tidy(juvenile_hare_cor)
```
```{r, include=FALSE}
#Call model diagnostics to check for normality 
plot(juvenile_hares_lm) # based on the qq plot, the residual distribution seems normal. The residuals, however, are not homoscedastic
```


```{r}
ggplot(data=juvenile_hares_date, aes(x=hindft, y=weight))+
  geom_point() +
  theme_classic() +
  labs(x= "Hind Foot Length (mm)",
       y= "Weight (g)", 
       title= "Juvenile Hare Hind Foot Length vs. Weight") +
  theme(plot.background = element_rect(fill="azure2")) +
  theme(panel.background = element_rect(fill="aliceblue")) +
  geom_smooth(method="lm", linetype= "dashed")
```

**Figure 3. Relationship between juvenile snowshoe hare weight and hind foot length** Individual observations are represented by points on the graph. The regression model is indicated by the dashed blue line which folows the equation: weight = `r round(juvenile_hares_lm_tidy$estimate[2], 2)`g mm ^-1^ * (hind foot length) + `r round(juvenile_hares_lm_tidy$estimate[1], 2)`g (p=`r round(juvenile_hares_lm_tidy$p.value[1], 2)`, R^2^ = `r round(juvenile_hares_lm_glance$r.squared[1], 3)` Pearson's *r* =`r round(juvenile_hares_cor_tidy$estimate,2)`) Data: Kielland et al. (2017)

There is a significant linear relationship in which juvenile hare weight can be predicted by hare juvenile hind foot length (p=`r round(juvenile_hares_lm_tidy$p.value[1], 2)`, R^2^ =`r round(juvenile_hares_lm_glance$r.squared[1], 3)`). The slope value of $\beta$ = `r round(juvenile_hares_lm_tidy$estimate[2], 2)` mm/g  indicates a predicted increase of `r round(juvenile_hares_lm_tidy$estimate[2], 2)` g in juvenile snowshoe hare weight for a 1 millimeter increase in hind foot length and there is a medium, positive, significant correlation between juvenile snowshoe hare hind foot length and weight (Pearson's *r* = `r round(juvenile_hares_cor_tidy$estimate,2)`, p= `r (juvenile_hares_cor_tidy$p.value)`). Undisplayed diagnostic plots indicate that the distribution of the residual is hetroscedastic and non-normal, which indicates that a linear model may not be the best fit.  


